---
title: "Exploring the Relationship between Fat Intake and Cholesterol Levels"
author: "Anna Nguyen, Allyson Ling , Bonnie Xu, Jennifer DeGuzman, Christine Le"
date: "May 10, 2020"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(purrr)
library(epitools)
library(lmtest)
library(biostat3)
```

# Load data
```{r}
questionnaire = read.csv("data/questionnaire.csv")
labs = read.csv("data/labs.csv")
exams = read.csv("data/examination.csv")
diet = read.csv("data/diet.csv")
demog = read.csv("data/demographic.csv")
data_list = list(questionnaire, labs, exams, diet, demog)
```

#Filter data to only inlude relevant variables, drop missing values
```{r}
questionnaire_vars = c("SEQN", "FSD032B", "FSD032C", "FSDHH")
labs_vars = c("SEQN", "LBXTC")
exams_vars = c("SEQN")
diet_vars = c("SEQN", "DR1TTFAT", "DR1TSFAT", "DR1TCARB")
demog_vars = c("SEQN", "DMDEDUC2", "RIDAGEYR", "RIAGENDR")
var_list = list(questionnaire_vars, labs_vars, exams_vars, diet_vars, demog_vars)

filtered_data = lapply(1:length(data_list), function(x) data_list[[x]] %>% dplyr::select(var_list[[x]]))
```

# Merge data frames
```{r}
full_data = filtered_data %>% reduce(full_join, by = "SEQN") %>% dplyr::select("SEQN", "DR1TTFAT", "LBXTC", "FSDHH") %>% drop_na()
colnames(full_data) = c("id", "fat_intake", "tot_chol", "food_security")
head(full_data) 

# Check how many observations there are
nrow(full_data)
```

-----------------------------------
Exposure = total daily fat intake
- Standardize using sample mean and SD
-----------------------------------

Transform X = full_data such that 

$$X^* = \frac{X - \mu_X}{\sigma_X}$$

```{r}
mu = mean(full_data$fat_intake)
sigma = sd(full_data$fat_intake)

full_data = full_data %>% mutate(fat_intake_transform = (fat_intake - mu) / sigma)
```


-----------------------------------
Outcome = cholesterol level
- High cholesterol level: tot_chol >= 240 mg/dL
- Normal cholesterol level: 160 <= tot_chol < 240mg/dL
- Low cholesterol level: tot_chol < 160 mg/dL (values dropped from data)
-----------------------------------

Let $Y$ represent cholesterol level such that 

$$Y=\begin{cases} 1, \text{if tot_chol} \geq \text{240 mg/dL} \\
                  0, \text{if tot_chol < 240 mg/dL} \end{cases}$$

```{r}
full_data = full_data %>% filter(tot_chol >= 160) %>%  
                          mutate(tot_chol = ifelse(tot_chol >= 240, 1, 0))
```

-----------------------------------
Confounding = household food security level
- Full security: 0
- Marginal security: 1
- Low security: 2
- Very low security: 3
-----------------------------------

Let $Z_i$ be an indicator of household food security level i

$$Z_0 = \begin{cases} 1, \text{if Full security} \\ 
                      0, \text{else} \end{cases}$$

$$Z_1 = \begin{cases} 1, \text{if Marginal security} \\ 
                      0, \text{else} \end{cases}$$
                     
$$Z_2 = \begin{cases} 1, \text{if Low security} \\ 
                      0, \text{else} \end{cases}$$

$$Z_3 = \begin{cases} 1, \text{if Very Low security} \\ 
                      0, \text{else} \end{cases}$$

Addionally, let $Z$ represent household food security level such that 

$$Z=\begin{cases} 0, \text{if Full security} \\
                  1, \text{if Marginal security} \\
                  2, \text{if Low security} \\
                  3, \text{if Very low security} \end{cases}$$
                  
```{r}
full_data = full_data %>% mutate(food_security = food_security-1)
```


# Model 1

$$log(\frac{p_x}{1-p_x}) = \alpha$$

```{r}
model1 = glm(formula = tot_chol ~ 1, 
             family = binomial(link='logit'), data = full_data)

summary(model1)
exp(model1$coefficients)
logLik(model1)
```

# Model 2

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X$$
- Where X = fat intake

```{r}
model2 = glm(formula = tot_chol ~ fat_intake, 
             family = binomial(link='logit'), data = full_data)

summary(model2)
exp(model2$coefficients)
logLik(model2)
```

P-value for fat variable is 0.663.

# Model 3

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X^*$$

```{r}
model3 = glm(formula = tot_chol ~ fat_intake_transform, 
             family = binomial(link='logit'), data = full_data)

summary(model3)
exp(model3$coefficients)
logLik(model3)
```

# Model 4

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X + \gamma_2 Z_2 + \gamma_3 Z_3 + \gamma_4 Z_4$$
- Using Z = 1 (Full security) as the reference level

```{r}
model4 = glm(formula = tot_chol ~ fat_intake + as.factor(food_security), 
             family = binomial(link='logit'), data = full_data)

summary(model4)
exp(model4$coefficients)
logLik(model4)
```

No feature has a significant p-value.


# Model 5

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X + \gamma Z$$

```{r}
model5= glm(formula = tot_chol ~ fat_intake + food_security, 
            family = binomial(link='logit'), data = full_data)

summary(model5)
exp(model5$coefficients)
logLik(model5)
```

No feature has a significant p-value.

# Model 6

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X + \gamma Z + \zeta (X \times Z)$$

```{r}
model6 = glm(formula = tot_chol ~ fat_intake + food_security + fat_intake:food_security, 
             family = binomial(link='logit'), data = full_data)

summary(model6)
exp(model6$coefficients)
logLik(model6)
lrtest(model6)
```

Interaction term has p-value of 0.0817.
Fat intake has p-value of 0.1988.
Food security has p-value of 0.1094.

LRT gives chi-square statistic of 3.1927 with a corresponding p-value of 0.3629.

# Model 7

$$log(\frac{p_x}{1-p_x}) = \alpha + \beta X + \gamma_1 Z_1 + \gamma_2 Z_2 + \gamma_3 Z_3 + \zeta_1(X \times Z_1) + \zeta_2(X \times Z_2) + \zeta_3(X \times Z_3) + \zeta_4(X \times Z_4)$$

```{r}
model7 = glm(formula = tot_chol ~ fat_intake + as.factor(food_security) + fat_intake:as.factor(food_security), 
             family = binomial(link='logit'), data = full_data)

summary(model7)
exp(model7$coefficients)
logLik(model7)
lrtest(model7)
```

All p-values are above 0.1. Fail to reject null-hypothesis for LRT.



