---
title: "work flow 2"
author: "Christine Le"
date: "5/1/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
title: "Exploring the Relationship between Fat Intake and Cholesterol Levels"
author: "Anna Nguyen, Allyson Ling , Bonnie Xu, Jennifer DeGuzman, Christine Le"
date: "May 10, 2020"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(purrr)
library(epitools)
library(lmtest)
library(biostat3)
library(broom)
```

# Load data
```{r}
questionnaire = read.csv("data/questionnaire.csv")
labs = read.csv("data/labs.csv")
exams = read.csv("data/examination.csv")
diet = read.csv("data/diet.csv")
demog = read.csv("data/demographic.csv")
data_list = list(questionnaire, labs, exams, diet, demog)
```

#Filter data to only inlude relevant variables, drop missing values
```{r}
questionnaire_vars = c("SEQN", "FSD032B", "FSD032C", "FSDHH")
labs_vars = c("SEQN", "LBXTC", "LBDLDL")
exams_vars = c("SEQN")
diet_vars = c("SEQN", "DR1TTFAT", "DR1TSFAT", "DR1TCARB")
demog_vars = c("SEQN", "DMDEDUC2", "RIDAGEYR", "RIAGENDR")
var_list = list(questionnaire_vars, labs_vars, exams_vars, diet_vars, demog_vars)

filtered_data = lapply(1:length(data_list), function(x) data_list[[x]] %>% dplyr::select(var_list[[x]]))
```

# Merge data frames
```{r}
full_data = filtered_data %>% reduce(full_join, by = "SEQN") %>% dplyr::select("SEQN", "DR1TTFAT", "LBXTC", "FSDHH", "RIDAGEYR", "RIAGENDR", "DR1TSFAT", "DR1TCARB") %>% drop_na()
colnames(full_data) = c("id", "fat_intake", "tot_chol", "food_security", "age", "gender", "saturated_fat", "carbohydrate")
head(full_data) 

# Check how many observations there are
nrow(full_data)
```


-----------------------------------
Outcome = cholesterol level
- High cholesterol level: tot_chol >= 240 mg/dL
- Normal cholesterol level: 160 <= tot_chol < 240mg/dL
- Low cholesterol level: tot_chol < 160 mg/dL (values dropped from data)
-----------------------------------

Let $Y$ represent cholesterol level such that 

$$Y=\begin{cases} 1, \text{if tot_chol} \geq \text{240 mg/dL} \\
                  0, \text{if tot_chol < 240 mg/dL} \end{cases}$$

```{r}
full_data = full_data %>% filter(tot_chol >= 160) %>%  
                          mutate(tot_chol = ifelse(tot_chol >= 240, 1, 0))
```

-----------------------------------
Confounding = household food security level
- Full security: 0
- Marginal security: 1
- Low security: 2
- Very low security: 3
-----------------------------------

Let $Z_i$ be an indicator of household food security level i

$$Z_0 = \begin{cases} 1, \text{if Full security} \\ 
                      0, \text{else} \end{cases}$$

$$Z_1 = \begin{cases} 1, \text{if Marginal security} \\ 
                      0, \text{else} \end{cases}$$
                     
$$Z_2 = \begin{cases} 1, \text{if Low security} \\ 
                      0, \text{else} \end{cases}$$

$$Z_3 = \begin{cases} 1, \text{if Very Low security} \\ 
                      0, \text{else} \end{cases}$$

Addionally, let $Z$ represent household food security level such that 

$$Z=\begin{cases} 0, \text{if Full security} \\
                  1, \text{if Marginal security} \\
                  2, \text{if Low security} \\
                  3, \text{if Very low security} \end{cases}$$
                  
```{r}
full_data = full_data %>% mutate(food_security = food_security-1)
```


-----------------------
Possible third variabe: Age

$$A = \text{age at interview, as a continuous variable} $$
----------------------

----------------------
Possible third variable : Gender
Self-reported Gender at Interview
$$ G = \begin{cases} 1, \text{Female} \\
                     0, \text{Male} \end{cases}
                     $$
```{r}
full_data <- full_data %>% mutate(gender = gender -1)
```

#Model 1

```{r}
model1 <- glm(tot_chol ~ fat_intake * as.factor(food_security) + fat_intake *age + fat_intake * gender + fat_intake *carbohydrate + age*gender, family = binomial(link = 'logit'), data = full_data)

summary(model1)
```

#Model 2 model without significant interaction
```{r}
model2 <- glm(tot_chol ~ fat_intake + age + gender + carbohydrate  + age*gender + fat_intake*carbohydrate, family = binomial(link = 'logit'), data = full_data)

summary(model2)
lrtest(model1, model2)
```

#Model 3 model without significant confounders

```{r}
model3 <- glm(tot_chol ~ fat_intake + age + gender + age*gender, family = binomial(link = 'logit'), data = full_data)

summary(model3)
lrtest(model2, model3)
```

